### 2022-11-13 chapter1.tex
Moved discussion of trees to results and re-wrote first paragraph of discussion

\subsection*{Inference of species trees}
% TODO: Add discussion of creating species tree

\begin{figure}[h!]
\includegraphics[width=\textwidth]{trees/out/trees1.png}
\centering
\caption{\textbf{Phylogenetic tree of species.}
Phylogenetic trees were fit to meta-alignments sampled from alignments created by the orthology inference pipeline with the additional genome of \textit{Scaptodrosophila lebanonensis} (sleb) as an outgroup. The trees fit to each meta-alignment were then combined into a single consensus tree. Values at nodes are bootstrap percentages. A) The LG model was fit to meta-alignments directly sampled from the original protein alignments B). The GTR model was fit to meta-alignments sampled from nucleotide alignments which were “reverse translated” from the original protein alignments and their corresponding coding sequences.}
\label{fig:trees}
\end{figure}

\section*{Discussion}
% TODO: Re-write to remove species tree and combine with following paragraph; also highlight specific contributions
The orthologous groups and alignments yielded by this pipeline are a valuable resource for comparative studies of gene birth/death processes and protein evolution at the level of both entire proteomes and specific gene families in the \textit{Drosophila} genus. To assist these efforts, consensus phylogenetic trees were inferred from meta-alignments sampled from alignments of single copy orthologous groups (Fig.~\ref{fig:trees}). Although this work focused on single copy orthologs, other studies may require different subsets of orthologous groups that demand other pre-processing and alignment strategies. Thus, the orthologous groups and the initial alignments with and without non-representative sequences are included as supplemental data.

The trends that permitted this work to substantially improve on previous efforts will quickly render its products obsolete. More \textit{Drosophila} genomes will be assembled and annotated, and the existing annotations will be updated. However, an authoritative and lasting set of orthologous groups and alignments is not the primary goal of this work. It instead serves as a case study in how dense taxonomic sampling and modern genome assembly and annotation pipelines present new opportunities and challenges to the traditional techniques for identifying and aligning orthologous groups.

@@

\subsection*{Inference of species trees}
Since the roots of phylogenetic trees are not identifiable with commonly used time-reversible substitution models, the orthology inference pipeline was repeated with the outgroup species \textit{Scaptodrosophila lebanonensis}. Orthologous groups with one sequence for each species were aligned, and 100 meta-alignments were constructed by randomly sampling 10,000 columns from these 9,435 alignments. (The alignments were not refined with HMMer before sampling.) To determine the effect of invariant columns and gaps, two sampling strategies were used where invariant columns were allowed or disallowed and the maximum fraction of gaps was set at 0, 50, and 100\%. Their combination yielded six different sets of meta-alignments. A tree was fit to each meta-alignment with the LG substitution model and four discrete gamma rate categories using IQ-TREE 1.6.12~\cite{Nguyen2014, Le2008, Yang1994}. If the sampling strategy allowed invariant columns, an invariant rate category was included. The resulting trees from each set were then merged into a majority consensus tree (Fig.~\ref{sfig:trees_LG}).

>>>>

\subsection*{Inference of species trees}
Many phylogenetic methods require a species tree to inform the evolutionary relationships between sequences. In fact, the phylo-HMMs discussed previously used a species tree as an input, though we omitted this detail for clarity of exposition. Therefore, to support the curation step and other downstream analyses, we sought to infer phylogenetic trees from the aligned sequences. However, since the roots of phylogenetic trees are not identifiable with commonly used time-reversible substitution models, we repeated the orthology inference pipeline with the outgroup species \textit{Scaptodrosophila lebanonensis}. Afterwards, we inferred phylogenetic trees using the LG model of amino acid substitution from 100 meta-alignments sampled from alignments of single copy orthologous groups. We then combined them into a single consensus tree (Fig.~\ref{fig:trees}A). To provide a similar tree for the analysis of non-coding regions, we inferred phylogenetic trees using the GTR model of nucleotide substitution from 100 meta-alignments sampled from nucleotide alignments which were "reverse translated" from the protein alignments and their corresponding coding sequences. The resulting consensus tree is shown in Figure~\ref{fig:trees}B. Both trees have an identical topology, which is consistent with other published phylogenies \cite{D12GC2007, DaLage2007}.

\begin{figure}[h!]
\includegraphics[width=\textwidth]{trees/out/trees1.png}
\centering
\caption{\textbf{Phylogenetic tree of species.}
A) The LG model was fit to meta-alignments directly sampled from the original protein alignments B). The GTR model was fit to meta-alignments sampled from "reverse translated" nucleotide alignments. Values at nodes are bootstrap percentages.}
\label{fig:trees}
\end{figure}

\section*{Discussion}
% TODO: Highlight specific contributions
The orthologous groups and alignments yielded by this pipeline are a valuable resource for comparative studies of gene birth/death processes and protein evolution at the level of both entire proteomes and specific gene families in the \textit{Drosophila} genus. To assist these efforts, the final alignments after curation and the labels from the "missing" phylo-HMM are provided as supplemental data. Although this work focused on single copy orthologs, other studies may require different subsets of orthologous groups that demand other pre-processing and alignment strategies. Therefore, we have included the orthologous groups and the initial alignments with and without non-representative sequences in the supplemental data as well. While we anticipate these resources will remain relevant in the near term, the trends that permitted this work to substantially improve on previous efforts will render them obsolete in the coming years as more \textit{Drosophila} genomes are assembled and annotated. However, an authoritative and lasting set of orthologous groups and alignments is not the primary goal of this work. Instead, it serves as a case study in how dense taxonomic sampling and modern genome assembly and annotation pipelines present new opportunities and challenges to the traditional techniques for identifying and aligning orthologous groups.

@@

\subsection*{Inference of species trees}
The orthology inference pipeline was first repeated with the outgroup species \textit{Scaptodrosophila lebanonensis}. Then orthologous groups with one sequence for each species were aligned, and 100 meta-alignments were constructed by randomly sampling 10,000 columns from these 9,435 alignments. (The alignments were not refined with HMMer before sampling.) To determine the effect of invariant columns and gaps, two sampling strategies were used where invariant columns were allowed or disallowed and the maximum fraction of gaps was set at 0, 50, and 100\%. Their combination yielded six different sets of meta-alignments. A tree was fit to each meta-alignment with the LG substitution model and four discrete gamma rate categories using IQ-TREE 1.6.12~\cite{Nguyen2014, Le2008, Yang1994}. If the sampling strategy allowed invariant columns, an invariant rate category was included. The resulting trees from each set were then merged into a majority consensus tree (Fig.~\ref{sfig:trees_LG}).
